### 2013-02-15 doc/icml2013/algo.tex

%%% second moments
Intuitively, performing regression on only $y$ given $x$ provides only first-order
information.  The key insight is that we can perform regression
on higher order powers to obtain more information about the parameters.
Specifically, for an integer $p \ge 1$, let us define the average
$m$-th order tensor power of the parameters as follows:
\begin{align}
M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
\end{align}
Now consider performing regression on $y^2$ given $x\tp{2}$.
Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
we have:
\begin{align}
y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h}{x} + (\epsilon^2 - \sigma^2). \nonumber
\end{align}
Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.\footnote{If $\sigma^2$ were not known,
we could treat it as another coefficient
to be estimated.  The coefficients $M_2$ and $\sigma^2$ can be estimated jointly
provided that $x$ does not already contain a bias ($x_j$ must be non-constant for every $j \in [d]$).}
Importantly, the noise has mean zero; 
in fact each of the three terms has mean zero
by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.

@@

In this section, we conduct a theoretical analysis of Spectral Experts.
In particular, we prove that it enjoys consistent parameter estimates.
\begin{theorem}[Convergence of Spectral Experts]
  \label{thm:convergence}
Each $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$
and that $d \ge k$.
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
Suppose the number of samples is
$n = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$ and
$n = \Omega\left( \sigma^6 L^6 R^12 k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$,
and that the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
$\Omega(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}})$,
then the parameter estimates $\hat\theta$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\pi_h^*}$.
\end{theorem}

>>>>

%%% second moments
Intuitively, performing regression on only $y$ given $x$ provides only first-order
information.  The key insight is that we can perform regression
on higher order powers to obtain more information about the parameters.
Specifically, for an integer $p \ge 1$, let us define the average
$p$-th order tensor power of the parameters as follows:
\begin{align}
M_p &\eqdef \sum_{h=1}^k \pi_h \beta_h\tp{p}. \label{eqn:Mp} % \in (\Re^{d})\tp{p}.
\end{align}
Now consider performing regression on $y^2$ given $x\tp{2}$.
Expanding $y^2 = (\innerp{\beta_h}{x} + \epsilon)^2$,
using the fact that $\innerp{\beta_h}{x}^p = \innerp{\beta_h\tp{p}}{x\tp{p}}$,
we have:
\begin{align}
y^2 &= \innerp{M_2}{x\tp{2}} + \sigma^2 + \eta_2(x), \label{eqn:y2} \\
\eta_2(x) &= \innerp{\beta_h\tp{2} - M_2}{x\tp{2}} + 2 \epsilon \innerp{\beta_h}{x} + (\epsilon^2 - \sigma^2). \nonumber
\end{align}
Again, we have expressed $y^2$ has a linear function of $x\tp{2}$
with regression coefficients $M_2$, plus a known bias $\sigma^2$ and noise.\footnote{If $\sigma^2$ were not known,
we could treat it as another coefficient
to be estimated.  The coefficients $M_2$ and $\sigma^2$ can be estimated jointly
provided that $x$ does not already contain a bias ($x_j$ must be non-constant for every $j \in [d]$).}
Importantly, the noise has mean zero; 
in fact each of the three terms has mean zero
by (i) definition of $M_2$, (ii) independence of $\epsilon$ and $h$,
and (iii) the fact that $\E[\epsilon^2] = \sigma^2$.

@@

In this section, we conduct a theoretical analysis of Spectral Experts.
In particular, we prove that it enjoys consistent parameter estimates.
\begin{theorem}[Convergence of Spectral Experts]
  \label{thm:convergence}
Each dataset $\sD_p$ (for $p = 1, 2, 3$) consists of $n$ i.i.d. points drawn from the mixture
of linear regressions model with parameter $\theta^*$
(having three independent copies simplifies the analysis).
Assume that $\E[\cvec(x\tp{p})\tp{2}] \succ 0$ for each $p \in \{1,2,3\}$
and that $d \ge k$.
Let $\|x\|_2 \le R$ with probability $1$
and let $\|\beta_h\|_2 \le L$ for all $h \in [k]$.
Suppose the number of samples is
$n = \Omega(d^p R^p \sigmamin(\Sigma_p)^{-1} \sqrt{p \log(d) \log(1/\delta)})$ and
$n = \Omega\left( \sigma^6 L^6 R^12 k \sigmamin(\Sigma_p)^{-2} \frac{\log^3 (1/\delta)}{\epsilon^2} \right)$,
and that the regularization strengths $\lambda_n^{(2)}$ and $\lambda_n^{(3)}$ are
$\Omega(\sigma^3 L^3 R^6 \sqrt{\frac{\log(1/\delta)}{n}})$,
then the parameter estimates $\hat\theta$ returned by
\algorithmref{algo:spectral-experts} (with the columns appropriate permuted)
satisfies 
$\|\hat\pi - \pi^*\|_{\infty} \le \epsilon$
and for all $h \in [k]$,
$\|\hat\beta_h - \beta^*_h\|_2 \le \frac{\epsilon}{\pi_h^*}$.
\end{theorem}
